---
title: "Example Report: KEY"
author: "Kirsten Bell Burdett"
institute: "Northwestern University: Feinberg School of Medicine"
date: "`r format(Sys.Date(),'%d %b %Y')`"
geometry: margin=2cm
output:
  pdf_document:
    toc: yes
    toc_depth: 2
header-includes: \usepackage{float, graphicx, amsmath, amssymb, tikz, caption}
  \usetikzlibrary{shapes,arrows,positioning,backgrounds}
---



```{r setup, include=FALSE, echo=FALSE}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE,
                      fig.path='figure/beamer-',
                        fig.align='center',
                        fig.show='hold',
                        size='tiny',
                        fig.width=7,
                        fig.height=5,
                        out.width='.9\\linewidth')

stdt <- date()
```


```{r echo=FALSE}
#============  Functions!!!  =============#

#=======get median and CI for survival outcomes  =========#
survmedCI <- function(dat, SurvOut, roundval){
  res <- surv_median(dat)
  medval <- round(res$median, roundval)
  CIval <- paste0("(",round(res$lower,roundval), " , ", round(res$upper,roundval), ")")
  MedianCI <- paste0(medval, " ", CIval)
  
  finalout <- tibble(SurvOut,MedianCI)  #make name so that you can distiguish multiple survival outcomes
  
  return(finalout = finalout)
}


#========== KM plot with dotted line at median and text for median ============#
kmplotfun <- function(dat, fitres, xlabel, titleval){
  gout <- ggsurvplot(fitres, data = dat, risk.table = "nrisk_cumevents", surv.median.line = "v",xlab = xlabel, ylab = "Probability", title = titleval) 
  
  gout$plot <- gout$plot+ 
                ggplot2::annotate("text", 
                                  x = .7*fitres$time[length(fitres$time)], y = 0.8, # x and y coordinates of the text
                                  label = paste0("Median = ",round(surv_median(fitres)$median,3), " Units") , size = 4)
  return(gout=gout)
}



#======= summary AEs listing by grade =========#
listeventsfun <- function(dat){
  limitvars <- dat %>% dplyr::select(subjid, Grade, Event, total_subj) 

  bygrade <- limitvars %>%
    group_by(Event, Grade, subjid) %>%
    dplyr::summarise(n()) %>%
    group_by(Event, Grade) %>%
    dplyr::summarise(n = n()) %>%
    dplyr::rename(`No.` = n) %>%
    ungroup()
  
  widedat <- spread(bygrade, Grade, `No.`)
  for (i in 2:dim(widedat)[2]){
    first <- widedat[i]
    widedat[i] <- replace(first, is.na(first), 0)
    names(widedat)[i] <- paste0("Grade ", names(widedat)[i])  }
  
  return(widedat = widedat)
}


#function to get column totals
makepretty <- function(dat, eventflag){
  datin <- dat[dat[,eventflag] =="Yes",]
  out1 <- listeventsfun(datin)
  
  newtest <- c(NA, colSums(out1[,2:dim(out1)[2]]))  #get col totals

  finalouta <- out1 %>% mutate_if(is.numeric, function(x) paste0(x, " (", round((x/unique(dat$total_subj)) * 100,2), "%)") )
  finalout <- rbind(newtest, finalouta)
  
  finalout[1,1] <- ""   #remove the NA place holder 
  
  for (i in 2:dim(finalout)[2]){
    first <- finalout[i]
    finalout[i] <- replace(first, first == "0 (0%)", "0")
  }
  return(finalout = finalout)
}
```

<!------------------------------------------------------------> 
<!------------------   Load Libraries  ----------------------->
<!------------------------------------------------------------>

```{r libraries, warning=FALSE, message=FALSE, echo=FALSE}
#Load libraries necessary for anlaysis. Set location of the data and output
library(survival)       # survival and veteran data
library(survminer)      # ggsurvplot      
library(tidyverse)
library(kableExtra)     # pretty kable output
library(tableone)       # CreateTableOne function

dir <- paste0(getwd(),"/Data/")         # path for data
outdir <- paste0(getwd(),"/Results/")   # path for exporting results (TLFs)
```


\vspace{15mm}

Data:
```{r message = FALSE}
readfile <- file.path(dir, "FakeExampleAEs.csv")
fakeAEdat <- read_csv(readfile)
tools::md5sum(readfile)
```
\vspace{1mm}
`r gsub(".*Data/","",readfile)` \
\vspace{1mm}
`r tools::md5sum(readfile)`



Veteran Data: \
```{r echo=FALSE}
#str(veteran)     #Lung Cancer Study

veteran2 <- veteran %>%
  dplyr::rename("Treatment" = trt, "Cell Type" = celltype, "Karnofsky Performance Score" = karno,
                "Age in Years"=age, "Prior Therapy" = prior) %>%
  dplyr::mutate(`Prior Therapy` = ifelse(`Prior Therapy` == 0, "No",
                                         ifelse(`Prior Therapy` == 10, "Yes", `Prior Therapy`)))


#=== we will only use the subjects with test trt and with cell type: squamous, smallcell, adeno

trt2only <- veteran2 %>% dplyr::filter(Treatment == 2)
#dim(veteran); dim(trt2only)

analysisdat <- trt2only %>% dplyr::filter(`Cell Type` %in% c("squamous", "smallcell", "adeno"))
#dim(analysisdat)
```
We will be using the veteran data from the survival package. The original data has `r nrow(veteran)` subjects, but we will be subsetting to only those subjects that have test treatment and cell type of squamous, smallcell, or adeno which will result in `r nrow(analysisdat)` subjects.
\vspace{10mm}




* * *



\pagebreak

<!------------------------------------------------------------> 
<!------------   Statistical Analysis Plan  ------------------>
<!------------------------------------------------------------>

<!--------  Section "Background"  --------------->
# Background
\underline{Protocol Title Here.} This is a great place to have background information about the study as well as include your **statistical analysis plan (SAP)**.

<!--------  Subsection "Objectives"  ------------>
##Objectives
\vspace{3mm}

\underline{Primary Objective:}
\begin{itemize}
	\item Define Here
\end{itemize}

\vspace{3mm}

\underline{Secondary Objective:}
\begin{itemize}
	\item Define Here
\end{itemize}


\vspace{20mm}


<!------------------------------------------------------------> 
<!-------------     CONSORT Flow Diagram    ------------------>
<!------------------------------------------------------------>
## CONSORT Flow Diagram

\vspace{5mm}

\tikzset{
  block_full/.style ={rectangle, draw=black, thick, fill=white,
    text width=20em,text centered, minimum height=4em, inner sep=6pt},
  block_noborder/.style ={rectangle, draw=none, thick, fill=none,
    text width=15em, text ragged, minimum height=1em, inner sep=6pt},
  block_arm/.style ={rectangle, draw=black, thick, fill=yellow,
  	text width=10em,text centered, minimum height=4em, inner sep=6pt},
  line/.style ={draw, very thin, -latex', shorten >=0pt}
  }
    
\sffamily
{\centering 
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.9}]   
  \node [block_full] (start) {
    \textbf{Overall Population}  \\
    \hspace{5pt} Total Patients (n= `r nrow(veteran2)`)}; 
    
    
  \node [block_full, below=of start] (testpop) {
    \textbf{Test Treatment Population:} \\
    Total Patients (n= `r nrow(trt2only)`) };
    
  \node [block_noborder, left=of testpop] (lefttest) {
    Excluded (n= `r nrow(veteran2) - nrow(trt2only)`): \\
    \hspace{5pt} Exclude Standard Treatment}; 
    
  \node [block_full, below=of testpop] (finalpop) {
    \textbf{Population of Interest:} \\
    Cell Type: Squamous, Small Cell, Adeno \\
    Total Patients (n=`r nrow(analysisdat)`)};
  
  \node [block_noborder, left=of finalpop] (excluded_large) {
    Excluded (n= `r nrow(trt2only)- nrow(analysisdat)`): \\
    \hspace{5pt} Excluded Cell Type: Large}; 
    
  \node [block_full, below=of finalpop] (finalbox) {
    \textbf{Final Box}};
    
  \node [block_arm, below=of finalbox, xshift = -2.5cm] (arma) {Arm A};
  
  \node [block_arm, right=of arma] (armb) {Arm B};

    
    % % Arrows % %
  \draw[->, line width=0.5pt] (start) -- (testpop);
  \draw[->, line width=0.5pt] (testpop) -- (finalpop);
  \draw[->, line width=0.5pt] (testpop) -- (lefttest);
  \draw[->, line width=0.5pt] (finalpop) -- (excluded_large);
  \draw[->, line width=0.5pt] (finalpop) -- (finalbox);
  \draw[->, line width=0.5pt] (finalbox) -- (arma);
  \draw[->, line width=0.5pt] (finalbox) -- (armb);
    
\end{tikzpicture}
}


\pagebreak


<!------------------------------------------------------------> 
<!---------------   Demographics with tableone --------------->
<!------------------------------------------------------------>
#Baseline Patient Demographic and Clinical Characteristics
```{r echo=FALSE, include = FALSE}
testtab <- print(CreateTableOne(vars = c("Cell Type", "Karnofsky Performance Score", "Age in Years", "Prior Therapy"), 
                                data = analysisdat, factorVars= c("Cell Type","Prior Therapy")), showAllLevels = TRUE)
```


<!----------------------------------------------------------------> 
<!-----------------   kable() and kableExtra()   ----------------->
<!-- #longtable=TRUE: include in case table spans multiple pages ->
<!-- #booktabs=TRUE: removes lines between each col and row      ->
<!-- #striped: rows alternate colors                             ->
<!-- #HOLD_position: hold position of table                      ->
<!-- #repeat_header: repeats header if table goes to next page   ->
<!---------------------------------------------------------------->

```{r results='asis'}
kable(testtab, longtable = TRUE, booktabs = TRUE, caption = "Demographics for Population of Interest") %>%
  kable_styling(latex_options = c("striped","HOLD_position", "repeat_header")) %>%
  footnote(general = "Footnote Here.", general_title = "")
```

```{r echo=FALSE}
write.csv(testtab, file.path(outdir, paste0("Demo_interest_veterans.csv" )))
```


\pagebreak

#Primary Objective
\underline{Primary Objective:} Define HERE


##Adverse Events

<!------------------------------------------------------------> 
<!---------------   Table with kableExtra   ------------------>
<!-- #row_spec(1, bold = TRUE ): makes the first row bold   -->
<!-- #column_spec(1, width = "15em"): change column 1 width -->
<!-- #threeparttable = T: for long footnotes                -->
<!------------------------------------------------------------>

```{r results='asis', message = FALSE}
kable(makepretty(fakeAEdat, "anyRAE"), longtable = TRUE, booktabs = TRUE, caption = "All Related Adverse Events by Grade: for each subject that had at least one event in the overall population.") %>%
  kable_styling(latex_options = c("striped", "repeat_header")) %>%
  row_spec(1, bold = TRUE ) %>%
  column_spec(1, width = "15em") %>%
  footnote(general = "Frequencies use the total number of subjects in the overall population as the denominator. The first row indicates event totals.", general_title = "", threeparttable = T)
```



\vspace{15mm}

##Survival Outcomes
```{r echo=FALSE}
#=========== Overall Population (veteran2)  ============#
allfit <- survfit(Surv(time, status) ~ 1, data = veteran2, type = "kaplan-meier")
allvals <- survmedCI(allfit, "OS - Overall Population", 3)
v1 <- kmplotfun(veteran2, allfit, "Overall Survival Time", "Overall Population")

# 6 and 12 unit survival probability
ALL6 <- summary(allfit,time = 6)
ALL12 <- summary(allfit,time = 12)



#========== Population of interest (analysisdat) =======#
vetfit <- survfit(Surv(time, status) ~ 1, data = analysisdat, type = "kaplan-meier")
vetvals <- survmedCI(vetfit, "OS - Population of Interest", 3)
v2 <- kmplotfun(analysisdat, vetfit, "Overall Survival Time", "Population of Interest")

# 6 and 12 unit survival probability
interest6 <- summary(vetfit,time = 6)
interest12 <- summary(vetfit,time = 12)


#=======   SHOW the Survival Outcome median(CI) =========#
showsurvout <- rbind(allvals, vetvals) 
showsurvout <- as_tibble(showsurvout) %>%
  dplyr::rename(`Population` = SurvOut, `Median (CI)` = MedianCI)
```

For population of interest (n=`r nrow(analysisdat)`) the 6 and 12 unit OS probability for the overall population was `r round(ALL6$surv,3)` and `r round(ALL12$surv,3)` respectively.  6 and 12 unit OS probability for the population of interest was `r round(interest6$surv,3)` and `r round(interest12$surv,3)` respectively.
```{r echo=FALSE, results='asis'}
kable(showsurvout, booktabs = TRUE) %>%
  kable_styling(latex_options = "striped")
```


<!------------------------------------------------------------> 
<!------------------   Kaplan-Meier Plots   ------------------>
<!------------------------------------------------------------>
```{r}
#Plot KM plots
allplots <- arrange_ggsurvplots(list(v1,v2), nrow=1, ncol=2, print=FALSE)
ggsave(paste0(outdir,"survplots.pdf"),allplots, width = 15, height = 8)
```

\vspace{5mm}

![Kaplan-Meier plots for overall survival.](Results/survplots.pdf)



\pagebreak



#Secondary Objective
\underline{Secondary Objective:} Define HERE



\pagebreak


<!------------------------------------------------------------> 
<!---------   Supplementary: Equation Example  --------------->
<!------------------------------------------------------------>
#Supplementary

####Survival Analysis Equation Example:

The Kaplan-Meier method is a non-parametric procedure which is sensitive to the choice of time interval and assumes that events of individuals are independent. We will denote the jth event time as $t_{(j)}$. The number at risk before $t_{(j)}$ is denoted as $n_{j}$ where $n_{j+1} = n_{j} - d_{j} - c_{j}$ where $c_{j}$ is the number of censored observations. The number of events at $t_{(j)}$ is denoted by $d_{j}$. The estimated survival function given below can be interpreted as the probability of survival through the interval $t_{(k)} \le t < t_{(k+1)}$ where k = 1,2...r where r is the event time.
	
\begin{align*}
\hat{S}(t) &= \prod^{k}_{j=1}\frac{n_{j} - d_{j}}{n_{j}}, && t_{(k)} \le t < t_{(k+1)}
\end{align*}
	
	
	The hazard is then estimated by taking the ratio of $d_{j}$ compared to $n_{j}\tau_{(j)}$ where $\tau_{(j)}$ is the length of the jth time interval ($t_{(j+1)} - t_{(j)}$). 
	
\begin{align*}
\hat{h}(t) &= \frac{d_{j}}{n_{j}\tau_{j}} , &&t_{(j)} \le t < t_{(j+1)}
\end{align*}



\pagebreak

<!------------------------------------------------------------> 
<!----------------   Session Information   ------------------->
<!------------------------------------------------------------>
Session Information
```{r echo=FALSE,results='asis'}
toLatex(sessionInfo(), locale=FALSE)
```
```{r echo=FALSE}
print(paste("Start Time",stdt))
print(paste("End Time  ",date()))
```




