---
title: "Word Example"
author: "Megan McCabe"
date: "`r format(Sys.Date(),'%d %b %Y')`"
output: word_document
---

```{r setup, include=FALSE}
library(broom)       # used for tidy()
library(tidyverse)
library(tableone)
library(survival)    # veteran data
library(knitr)

stdt <- date()
```

# Linear Regression Model
```{r}
fit <- lm(dist ~ speed, data = cars)
fit_res <- tidy(fit); fit_res
```

Easily pull out results with tidy() function. Then we can use inline code like `format.pval(fit_res$p.value[fit_res$term == "speed"], digit = 4)` to show the p-value  `r format.pval(fit_res$p.value[fit_res$term == "speed"], digit = 4)` corresponding to speed. \






# Baseline Patient Demographic and Clinical Characteristics
```{r echo=FALSE, include = FALSE}
veteran2 <- veteran %>%
  dplyr::rename("Treatment" = trt, "Cell Type" = celltype, "Karnofsky Performance Score" = karno,
                "Age in Years"=age, "Prior Therapy" = prior) %>%
  dplyr::mutate(`Prior Therapy` = ifelse(`Prior Therapy` == 0, "No",
                                         ifelse(`Prior Therapy` == 10, "Yes", `Prior Therapy`)))

tabout <- print(CreateTableOne(vars = c("Cell Type", "Karnofsky Performance Score", "Age in Years"),
               data = veteran2, strata = "Prior Therapy", test = FALSE), showAllLevels = TRUE)
```

```{r echo=FALSE}
kable(tabout, caption = "Demographics by Prior Therapy") 
```


## Specifying Reported Summary Statistic

Depending on the distribution of a continuous variable, you may want to report a different summary statistic in your Table 1. For instance, we can look at the distribution of the age variable.

```{r echo=FALSE, fig.width=4.5, fig.height=3.5}

ggplot(veteran, aes(age)) + 
  geom_histogram(bins = 20) +
  theme_bw() +
  ggtitle("Distribution of Age (years)") +
  xlab("Age (years)") +
  ylab("Count")

```

Based on the histogram, we may want to report the median (IQR) instead of the mean (sd) as reported earlier. Below we will recreate the table with that change.
 
```{r echo=FALSE, include = FALSE}

tabout2 <- print(CreateTableOne(vars = c("Cell Type", "Karnofsky Performance Score", "Age in Years"),
               data = veteran2, strata = "Prior Therapy", test = FALSE), showAllLevels = TRUE, nonnormal = c("Age in Years"))
```

```{r echo=FALSE}
kable(tabout2, caption = "Demographics by Prior Therapy") 
```



# Session Information
```{r echo=FALSE}
sessionInfo()
print(paste("Start Time",stdt))
print(paste("End Time  ",date()))
```




